---
title: "CDem: EDA"
author: "Rainer"
output: 
  html_document: 
    toc: true
---

## imports
```{r, message=FALSE}
library(dplyr)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(corrplot)

Cleaned_CES_2021_Data <- read_csv("numerical_Cleaned_CES_2021_Data.csv")
```
## how does age correlate to politics  

  lets investigate some interesting correlations. year of birth seems to be consistently negatively correlated with new consumption, political interest, likelihood to vote, populism etc. lets do some quick visual  
  
  <br>
  
  news consumption vs general interest in politics, coloured by year of birth
```{r}
Cleaned_CES_2021_Data |>
  ggplot(aes(x = cps21_news_cons, y = cps21_interest_gen_1, colour = cps21_yob)) +
  geom_jitter(width = 0.5, height = 0.5)
```
<br>

   general interest vs year of birth
  news con vs year of birth
```{r}
Cleaned_CES_2021_Data |>
  ggplot(aes(x = cps21_interest_gen_1, y = cps21_yob)) +
  geom_jitter(width = 0.5, height = 0.5)

Cleaned_CES_2021_Data |>
  ggplot(aes(x = cps21_news_cons, y = cps21_yob)) +
  geom_jitter(width = 0.5, height = 0.5)
```

lets check numerically how much its actually correalted w a linear regresion model4


















  news consumption vs general interest in politics, coloured by political partisanship
```{r}
Cleaned_CES_2021_Data |>
  ggplot(aes(x = cps21_news_cons, y = cps21_interest_gen_1, colour = pes21_pidtrad)) +
  geom_jitter(width = 0.5, height = 0.5, size = 0.2) +
  facet_wrap(vars(pes21_pidtrad), nrow = 4)
```
<br>
funky. partisonship has nothing to do with political interest/news consumption, even thought younger people have been well documented to be generally more liberal. lets check this  
```{r}
Cleaned_CES_2021_Data |>
  filter(pes21_pidtrad == c("conservative","liberal","ndp","green","bloc québécois"))|>
  ggplot(aes(x = cps21_yob, fill = pes21_pidtrad, colour = pes21_pidtrad)) +
  geom_bar(position = "fill", width = 3)
```
 
<br>
yup. overwhelmingly liberal and, surprisingly, a lot of people born in the 1990 range vote ndp. 
<br>

anyways, quick summary: age is correlated to political interest and partisanship. but political interest and partisanship do not interact. age is an *independent* predictor of partisanship.

# question proposal: what are the main predictors of partisanship?
how should we approach this?    

- make a lot of visuals and see if anything comes up. this is technically feasible, but a lot of effort. also not very scalable. skip for now    

- multi linear regression is not viable. partisanship is a cotegorical variable  

- ML model seems to be a good bet to me... 


